---
title: "Motor-imagery decoding with BCIFeatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Motor-imagery decoding with BCIFeatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(1)
```

BCIFeatR gives every feature family the **same** train/test interface, so you can
swap feature methods and classifiers without rewriting your pipeline. This
vignette walks through a complete, leakage-safe motor-imagery (MI) decoding run
on simulated EEG.

```{r setup}
library(BCIFeatR)
```

## Data format

A dataset is a **list of trial matrices** (`samples x channels`, same number of
channels in every trial) plus a **factor of labels**, one per trial.

```{r simulate}
# Simulate a 2-class MI dataset with band-power lateralization on two "motor"
# channels (a stand-in for C3/C4 mu/beta desynchronization).
simulate_mi <- function(n_per_class, n_ch = 8L, n_time = 256L, fs = 128, seed = 1) {
  set.seed(seed)
  y <- factor(rep(c("left", "right"), each = n_per_class))
  tvec <- seq(0, (n_time - 1) / fs, length.out = n_time)
  x <- lapply(seq_along(y), function(i) {
    M <- matrix(rnorm(n_time * n_ch, sd = 0.7), n_time, n_ch)
    a <- if (y[i] == "left") c(1.1, 0.4) else c(0.4, 1.1)   # swap by class
    M[, 3] <- M[, 3] + a[1] * sin(2 * pi * 10 * tvec)        # 10 Hz on ch 3
    M[, 6] <- M[, 6] + a[2] * sin(2 * pi * 22 * tvec)        # 22 Hz on ch 6
    M
  })
  list(x = x, y = y, fs = fs)
}

train <- simulate_mi(40, seed = 1)
test  <- simulate_mi(40, seed = 2)
length(train$x); dim(train$x[[1]]); table(train$y)
```

## One API, many feature families

`featEx4Train()` fits any feature extractor on the training trials and returns
the feature matrix plus a fitted `object`; `featEx4Test()` applies that exact
object to new trials (all train-time state — CSP filters, tangent reference,
preprocessing stats — is reused, never re-estimated on test).

```{r features}
fit  <- featEx4Train(train$x, train$y, feature = "CSP", params = list(ncomps = 4L))
Xtr  <- fit$features                       # trials x features
Xte  <- featEx4Test(test$x, fit$object, feature = "CSP")
dim(Xtr); dim(Xte)
```

Swapping the feature family is a one-word change. A small helper extracts any
method and reports its dimensionality:

```{r feature-zoo}
specs <- list(
  logvar     = list(),
  bandpower  = list(fs = 128, frequency_bands = list(c(8, 13), c(13, 30))),
  Hjorth     = list(),
  CSP        = list(ncomps = 4L),
  FBCSP      = list(fs = 128, frequency_bands = list(c(8, 13), c(18, 26)), ncomps = 2L),
  TS         = list(cov_type = "oas"),
  Riemannian = list(cov_type = "oas")
)
sapply(names(specs), function(f)
  ncol(featEx4Train(train$x, train$y, f, specs[[f]])$features))
```

## A classifier and accuracy

The package ships lightweight classifiers with a uniform `predict()` interface.
Here is the Riemannian Minimum Distance to Mean (MDM), which works directly on
raw trials:

```{r mdm}
m <- mdm_train(train$x, train$y, metric = "logeuclid")
m                                          # one-line model summary (print method)
pred <- predict(m, test$x)                 # class factor
mean(pred == test$y)                       # accuracy
```

Or pair any extracted feature with a downstream classifier:

```{r elastic-net}
fit <- featEx4Train(train$x, train$y, "TS", list(cov_type = "oas"))
cl  <- multiclass_EL(fit$features, train$y, alpha = 0.5, lambda = 0.1)
acc <- mean(cl$predict(featEx4Test(test$x, fit$object, "TS")) == test$y)
acc
```

## Preprocessing and attention

Channel preprocessing (`none`/`center`/`scale`/`whiten`) and optional
parameter-free SimAM attention re-weighting are applied **before** any feature,
through `params`:

```{r preprocess}
fit <- featEx4Train(train$x, train$y, "TS",
                    params = list(cov_type = "oas",
                                  preprocess = "whiten",
                                  simam = "vanilla"))
dim(fit$features)
```

## Fusing feature views with MCCA

When several feature families capture complementary information, multiset CCA
aligns them into a shared low-dimensional representation:

```{r mcca}
V1 <- featEx4Train(train$x, train$y, "logvar", list())$features
V2 <- featEx4Train(train$x, train$y, "bandpower",
                   list(fs = 128, frequency_bands = list(c(8, 13), c(13, 30))))$features
mc <- mcca_train(list(V1, V2), ncomp = 3L)
mc
dim(mcca_transform(mc, list(V1, V2))$consensus)
```

## Cross-session input

When you have multiple recording sessions, pass a list of trial-lists and a list
of label-vectors; each session is fitted/transformed independently and returned
as a list of per-session results.

```{r sessions}
multi <- featEx4Train(
  x = list(train$x, test$x),
  y = list(train$y, test$y),
  feature = "logvar", params = list()
)
length(multi); dim(multi[[1]]$features)
```

## Where to go next

* `?featEx4Train` lists every supported `feature` and its `params`.
* `mdm_train()` / `msvar_train()` are generative classifiers; `mcca_train()`
  fuses views; `simAM()` is the standalone attention transform.
* The Markov-switching VAR feature/classifier (`feature = "MSVAR"`,
  `msvar_train()`) is strongest on low-dimensional (e.g. CSP-projected) trials.
